\fullpagehumor[{Radio Yerevan Jokes}]{

    \textbf{Q: Radio Yerevan was asked: "Capitalism is the exploitation of man by man. What about Socialism?"} \\  
    \textbf{A:} Radio Yerevan answered: "Under Socialism it is exactly the other way around."  

    \medskip  
    \textbf{Q: Is there freedom of speech in the Soviet Union the same as in the USA?} \\  
    \textbf{A:} In principle, yes. In the USA, you can stand in front of the Washington Monument and yell, "Down with Reagan!" and you will not be punished. In the Soviet Union, you can stand in Red Square and yell, "Down with Reagan!" and you will not be punished.  

    \medskip  
    \textbf{Q: What is chaos?} \\  
    \textbf{A:} We do not comment on national economics.  

    \medskip  
    \textbf{Q: Is it true that half of the members of the Central Committee are fools?} \\  
    \textbf{A:} What a crazy question. Half of the members of the Central Committee are not fools!  

    \medskip  
    \textbf{Q: Is it true that Ivan Ivanovich Ivanov from Moscow won a car in a lottery?} \\  
    \textbf{A:} In principle, yes. But: it wasn't Ivan Ivanovich Ivanov but Aleksander Aleksandrovich Aleksandrov; he is not from Moscow but from Odessa; it was not a car but a bicycle; and he didn’t win it — it was stolen from him.  

    \medskip  
    \textbf{Q: Why is our government not in a hurry to land our men on the moon?} \\  
    \textbf{A:} What if they refuse to return?  

    \medskip  
    \textbf{Q: We are told that communism is already visible on the horizon. What then is a horizon?} \\  
    \textbf{A:} An imaginary line that moves farther away each time you approach it.  

    \medskip  
    \textbf{Q: Is it true that conditions in our labor camps are excellent?} \\  
    \textbf{A:} In principle, yes. Five years ago, one of our listeners was skeptical, so he was sent to investigate. He must have liked it so much that he hasn’t returned yet.  

    \medskip  
    \textbf{Q: What is the most beautiful city in the Soviet Union?} \\  
    \textbf{A:} Yerevan, obviously.\\
    \textbf{Q: How many nuclear bombs will it take to destroy Yerevan?} \\  
    \textbf{A:} Baku is also a beautiful city.  

    \medskip  
    \textbf{Q: Is it true that the United States is on the edge of a precipice?} \\  
    \textbf{A:} True, but we are a step farther than them.  

    \medskip  
    \medskip  
    \textbf{Q: Where do children grow up healthy, happy, and confident about their bright and wonderful future?} \\  
    \textbf{A:} In the Soviet Union!  
    \\(At this moment, cheering is interrupted by the sound of a child crying.)\\
    \textbf{Q: Verochka, why are you crying?} \\  
    \textbf{A:} I want to live in the Soviet Union...  

}
