\begin{technical}
{\Large\textbf{Micromanipulation and Mitochondrial Heteroplasmy}}\\[0.3em]

\noindent\textbf{Technical Challenges in Nuclear Transfer}

Maternal spindle transfer requires extracting the metaphase II spindle-chromosome complex without disrupting chromosome alignment. The spindle, a 15–20 μm birefringent structure visible under polarized light (Oosight™), must be aspirated with minimal surrounding cytoplasm. Tachibana et al. (2009) demonstrated this technique in rhesus macaques, achieving no detectable donor mtDNA carryover and producing healthy, fertile offspring. In human oocytes, Tachibana et al. (2013) showed that spindle transfer could be performed with minimal mitochondrial carry-over.

Critical technical parameters:
\begin{itemize}[leftmargin=*]
\item \textbf{Timing}: Within 2 hours of oocyte retrieval before spindle depolymerization
\item \textbf{Pipette diameter}: 20 μm beveled, approaching at 30° to minimize membrane deformation
\item \textbf{Cytoplasmic volume}: <5 picoliters co-aspirated to keep mitochondrial carry-over below 2\%
\item \textbf{Fusion method}: Electrofusion (1.0 kV/cm, 50 μs pulses) or HVJ-E (Sendai virus extract)
\end{itemize}

Pronuclear transfer exploits the 8–10 hour window post-fertilization when pronuclei are visible but unfused. Both pronuclei must be extracted together within a karyoplast — a membrane-bound cytoplasmic package containing ~5\% of oocyte volume. This preserves their relative positioning, which carries epigenetic information essential for proper development.

\noindent\textbf{Mitochondrial Carry-over and Detection}

Even stringent micromanipulation cannot eliminate all donor mitochondria. Reported carry-over levels in embryo reconstruction studies include:
\begin{itemize}[leftmargin=*]
\item PNT: typically <2\%, occasionally higher (Hyslop et al., 2016)
\item MST: approximately 1–2\% on average (Tachibana et al., 2013)
\item PB1T: often <0.5\% when optimized (reported in polar body transfer studies)
\end{itemize}

Deep sequencing can achieve ~0.1\% sensitivity, with digital droplet PCR detecting heteroplasmy as low as ~0.01\% in optimized assays.

\noindent\textbf{Heteroplasmy Dynamics in Development}

Low-level heteroplasmy behaves unpredictably during development due to:

\textbf{The mitochondrial bottleneck}: During oogenesis, mtDNA copy number drops from ~100,000 in mature oocytes to ~200 in primordial germ cells before clonal expansion. This bottleneck allows random genetic drift to dramatically shift heteroplasmy levels between generations.

\textbf{Tissue-specific segregation}: Post-mitotic tissues show divergent heteroplasmy patterns:
\begin{itemize}[leftmargin=*]
\item \textbf{Muscle}: Can amplify from 1\% to 50\% by adulthood
\item \textbf{Blood}: Typically maintains stable levels
\item \textbf{Brain}: Shows regional variation, with high-energy regions (substantia nigra) potentially enriching for wild-type mtDNA
\end{itemize}


\noindent\textbf{Nuclear-Mitochondrial Compatibility}

The mitochondrial proteome comprises ~1,500 proteins: 13 encoded by mtDNA, the remainder nuclear-encoded and imported. OXPHOS complexes require precise stoichiometry between nuclear and mitochondrial subunits.

Potential incompatibilities arise where mtDNA-encoded subunits must interact with nuclear-encoded partners (e.g., Complex I: 7 mitochondrial, 38 nuclear subunits). However, the Tachibana primate studies showed reassuring intraspecies compatibility, with healthy offspring produced using rhesus macaque nuclear-cytoplasmic combinations.

\vspace{0.5em}
\noindent\textbf{References:}\\
{\footnotesize
Tachibana, M. \textit{et al.} (2009). Mitochondrial gene replacement in primate offspring and embryonic stem cells. \textit{Nature} \textbf{461}, 367–372.\\
Tachibana, M. \textit{et al.} (2013). Towards germline gene therapy of inherited mitochondrial diseases. \textit{Nature} \textbf{493}, 627–631.
}
\end{technical}