\fullpagehumor[The Conductor]{
For years, the conductor had been great at his job. When the train arrived at the station, he'd blow his whistle to announce the arrival and the opening of the doors. After everyone got off and on, he'd blow the whistle again for the doors to close and the train to depart.

\medskip
It was a routine job, and one day while working — blowing his whistle as trains came and went — he thought everyone had boarded and blew his whistle. Unfortunately, an elderly passenger was still boarding. The doors closed with tragic consequences.

\medskip
Found responsible for this tragedy, he was tried in court and found guilty of manslaughter. The sentence was death by electric chair. A few hours before his execution, the warden visited his cell.

\medskip
"You get one last meal, anything you want," said the warden.

\medskip
"Really?" the man replied. "Well, if I were to have one last meal, it would be 5kg of bananas."

\medskip
The warden was stunned but arranged for the unusual request. He watched in amazement as the man devoured the bananas.

\medskip
Afterward, the man was led to the electric chair, strapped in, and connected to the electrodes. The executioner counted down: "3... 2... 1..." He pulled the lever — but nothing happened. Multiple attempts yielded the same result.

\medskip
"This can only be an act of God," the warden declared. "You're free to go."

\medskip
The man left and, remarkably, got another job as a train conductor. Eventually, another accident occurred when he blew his whistle while a child was retrieving a dropped toy from the platform.

\medskip
Again found guilty, he was sent to prison. The warden visited once more.

\medskip
"Let me guess, 5kg of bananas?" the warden asked.

\medskip
"Yes, please," said the man.

\medskip
Again he ate the bananas, again the execution failed, and again he was released.

\medskip
Incredibly, he returned to his conductor job a third time. Despite being careful for months, another accident eventually occurred.

\medskip
Sentenced to death for the third time, he again requested his 5kg of bananas. When the execution failed yet again, the warden couldn't contain his curiosity.

\medskip
"I have to know," he asked. "How are the bananas protecting you?"

\medskip
The man looked up and shrugged. "The bananas had nothing to do with it. I'm just a bad conductor."
}