\fullpagehumor[Time is Relative]{
A guy’s car breaks down on a rural road. No cell phone reception. Fortunately, there’s a farmhouse nearby, so he walks over to ask for help.

He rings the doorbell — no answer. But he hears some rustling from around the side of the house. He walks over and finds a farmer next to a pen with three ferrets.

“Excuse me — ” the guy begins, but the farmer cuts in: “Hold on, I have to feed the ferrets. I’ll be with you when I’m done.”

\medskip

The guy watches as the farmer picks up one of the ferrets, carries it to an apple tree with a ladder leaning against it, climbs the ladder while holding the ferret, lifts it up to an apple — *chomp* — then climbs back down and returns the ferret to the pen.

“Sorry, I just — ” the guy tries again, but the farmer grabs the second ferret. “Almost done.”

Same thing: apple tree, ladder, apple, chomp, back down, back to the pen.

\medskip

Then the third ferret. The farmer lifts it gently, murmurs something to it, and makes the familiar journey. Step by step up the ladder, hoists the ferret to a fresh apple — another bite. He climbs down slowly, cradles the ferret all the way back, and nestles it into the pen like it’s a newborn.

\medskip

Finally, the farmer turns. “Now — how can I help you?”

The guy says, “My car broke down and I need to call AAA. But first... wouldn't it be a lot faster to just pick the apples and toss them into the pen?”

The farmer pauses, thinking. Then nods.

\medskip

“Yeah... I suppose it would be faster that way.”

\medskip

He shrugs.

“But what’s time to a ferret?”
}
