\begin{historical}
Pirkei Avot opens with the chain of tradition: "Moses received the Torah from Sinai and transmitted it to Joshua, and Joshua to the Elders, and the Elders to the Prophets, and the Prophets transmitted it to the Men of the Great Assembly."

This transmission of authority defined who could determine Jewish law, including calendar matters. The chain continued through specific named authorities: Shimon the Righteous (one of the last of the Great Assembly), Antigonus of Socho, then paired leaders through the generations — the Zugot (pairs), where one served as Nasi (president) and one as Av Beit Din (head of the court).

The pairs included Yose ben Yoezer and Yose ben Yochanan, Joshua ben Perachya and Nittai of Arbel, Judah ben Tabbai and Shimon ben Shetach, Shemaya and Avtalyon, and finally Hillel and Shammai. From Hillel descended a dynasty of leaders who held the title of Nasi through the destruction of the Second Temple in 70 CE and beyond.

During the Temple period, this leadership controlled calendar determination. The Sanhedrin, with the Nasi presiding, declared new months based on witness testimony and intercalated years to maintain seasonal alignment. Their authority to declare time derived from the biblical verse "These are the appointed seasons of the Lord, which you shall proclaim" — the Hebrew emphasizes "which YOU shall proclaim," granting human authorities the power to establish sacred time.

After the Temple's destruction in 70 CE, the Sanhedrin reconvened in Yavneh under Rabban Yochanan ben Zakkai, then moved through various Galilean cities: Usha, Shefar'am, Beit She'arim, Sepphoris, and finally Tiberias. Despite lacking a Temple, they maintained calendar authority through the traditional chain of ordination (semicha) that connected each generation back to Moses.

The Roman Empire increasingly restricted Jewish self-governance. Emperor Hadrian outlawed ordination after the Bar Kokhba revolt (132-135 CE). Constantine I (306-337 CE) further limited Jewish courts' jurisdiction.

Hillel II served as Nasi from approximately 320 to 385 CE. Facing intensifying persecution and the imminent collapse of centralized Jewish authority, he made an unprecedented decision around 358 CE: publish the mathematical secrets of calendar calculation.

The rules Hillel II revealed included:

The precise length of the lunar month: $29$ days, $12$ hours, $44$ minutes, $3\tfrac{1}{3}$ seconds

The 19-year Metonic cycle with leap years in years $3$, $6$, $8$, $11$, $14$, $17$, and $19$

Four postponement rules (dechiyot) preventing Rosh Hashanah from falling on Sunday, Wednesday, or Friday:
- Lo ADU: Direct postponement from these days
- Molad Zaken: If the molad occurs after noon
- GaTRaD: In regular years, if molad falls on Tuesday after $9$ hours, $204$ parts
- BeTuTaKPaT: After leap years, if molad falls on Monday after $15$ hours, $589$ parts

These rules ensure Yom Kippur never falls adjacent to Shabbat (avoiding two consecutive days of severe restrictions) and that Hoshana Rabbah never falls on Shabbat (preserving the willow-beating ritual).

Hillel II's publication transformed calendar authority from political power to mathematical algorithm. Communities worldwide could now calculate Jewish dates independently, maintaining unity despite geographical dispersion. The precision apparent in these rules reflected centuries of astronomical observation and rabbinical refinement.

The chain of authority from Moses through the prophets, the Great Assembly, the Zugot, and the Nesi'im (plural of Nasi) had preserved and developed this knowledge. When centralized authority became unsustainable, Hillel II ensured continuity by making the secret knowledge public. His calendar remains the foundation of Jewish temporal practice today, unchanged for over 1,600 years.

This transition from human declaration to mathematical calculation represents one of Judaism's most successful adaptations to historical necessity. The authority to proclaim sacred time, transmitted through generations of leaders, became encoded in an algorithm accessible to any learned Jew. Hillel II's decision preserved not just the calendar but the unity of Jewish practice across continents and centuries.
\end{historical}